---
title: "Final Project"
author: "Kathryn Ellestad"
date: "`r Sys.Date()`"
output:
  pdf_document: default
  html_document:
    df_print: paged
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(
  echo = TRUE,
  fig.width = 8, 
  fig.height = 8)
```

## Libraries

```{r}
library(readxl)
library(ggplot2)
library(tidyverse)
library(plotly)
```



# Aesthetics

```{r}
if (knitr::is_latex_output()) {
  # PDF output: light theme
  theme_set(
    theme_minimal(base_size = 14, base_family = "Georgia") +
      theme(
        plot.title = element_text(face = "bold", color = "black", hjust = 0.5),
        plot.subtitle = element_text(color = "black", hjust = 0.5),
        axis.title.x = element_text(color = "black", margin = margin(t = 10)),
        axis.title.y = element_text(color = "black", margin = margin(r = 10)),
        axis.text.x = element_text(color = "black"),
        axis.text.y = element_text(color = "black"),
        panel.background = element_rect(fill = "white", color = NA),
        plot.background = element_rect(fill = "white", color = NA),
        panel.grid.major.x = element_line(color = "grey70"),
        panel.grid.major.y = element_line(color = "grey70"),
        panel.grid.minor = element_blank(),
        legend.position = "none"
      )
  )
} else {
  # HTML output: dark theme
  theme_set(
    theme_minimal(base_size = 14, base_family = "Georgia") +
      theme(
        plot.title = element_text(face = "bold", color = "#f8f8f8", hjust = 0.5),
        plot.subtitle = element_text(color = "#cccccc", hjust = 0.5),
        axis.title.x = element_text(color = "#f8f8f8", margin = margin(t = 10)),
        axis.title.y = element_text(color = "#f8f8f8", margin = margin(r = 10)),
        axis.text.x = element_text(color = "#f8f8f8"),
        axis.text.y = element_text(color = "#f8f8f8"),
        panel.background = element_rect(fill = "#222222", color = NA),
        plot.background = element_rect(fill = "#222222", color = NA),
        panel.grid.major.x = element_line(color = "grey30"),
        panel.grid.major.y = element_line(color = "grey30"),
        panel.grid.minor = element_blank(),
        legend.position = "none"
      )
  )
}

pastel_colors <- c("Physical" = "#C3E2C2", "Eink" = "#D8C7FF")

#used https://ggplot2.tidyverse.org/reference/ggtheme.html to learn how to change the theme.
```

## Data

```{r}
SleepandReadData <- read_excel("C:/Users/kathr/OneDrive/School/Data211/Final Project/Data 211 project Tracking.xlsx")
SleepandReadData <- SleepandReadData %>%
  mutate(
    Format = str_to_title(Format))

```
## Hypothesis 1: Does the format I use significantly affect the minutes spent reading?

### Hypothesis

$\mu_e =$ Average minutes spent reading eink

$\mu_p =$ Average minutes spent reading physical

$H_0: \mu_e = \mu_p$ 

$H_a: \mu_e \neq \mu_p$

```{r}
t.test(MinutesReading ~ Format, data = SleepandReadData, alternative = "two.sided")

```

**Conclusion:**

p-value = .4004 > $\alpha = .05$, therefore with a $5\%$ significance level, we fail to reject $H_0$. There is not sufficient evidence to support the claim that the format used significantly affects the amount of minutes spent reading.

```{r}

Boxplot <- ggplot(SleepandReadData, aes(x = Format, y = MinutesReading, fill = Format)) +
  geom_boxplot(width = 0.6, 
               outlier.shape = 21, 
               outlier.fill = "white", 
               outlier.color = "black", 
               outlier.size = 2) +
  scale_fill_manual(values = pastel_colors) +
  labs(
    title = "Reading Time by Format",
    y = "Minutes Reading",
    x = "Format"
  ) +
  coord_flip()

ggplotly(Boxplot)

```

## Hypothesis 2: Is the amount of sleep I get different based on the format I read before bed?

### Hypothesis

$\mu_e =$ Average sleep minutes after eink

$\mu_p =$ Average sleep minutes after physical

$H_0: \mu_e = \mu_p$ 

$H_a: \mu_e \neq \mu_p$

```{r}
t.test(TotalTimeAsleep ~ Format, data = SleepandReadData, alternative = "two.sided")
```

**Conclusion:**

p-value = .09576 > $\alpha = .05$, therefore with a $5\%$ significance level, we fail to reject $H_0$. There is not sufficient evidence to support the claim that the format chosen greatly affects the amount of sleep I get each night.

```{r}
SummaryTable <- SleepandReadData %>%
  group_by(Format) %>%
  summarise(
    ave = mean(TotalTimeAsleep, na.rm = TRUE),
    se = sd(TotalTimeAsleep, na.rm = TRUE) / sqrt(n()),
    n = n(),
    tstar = qt(1 - 0.05/2, df = n() - 1)
  )

SleepvsFormat <- ggplot(SummaryTable, aes(x = Format, y = ave, fill = Format)) +
  geom_col(width = 0.5, 
           color = "white") +
  geom_errorbar(aes(ymin = ave - tstar * se, ymax = ave + tstar * se), 
                width = 0.2, 
                linewidth = 0.8) +
  geom_point(shape = 21, 
             size = 4, 
             color = "black", 
             fill = "white", 
             stroke = 1.2) +
  scale_fill_manual(values = pastel_colors) +
  labs(title = "Average Total Sleep by Format",
       y="Average Minutes Slept")

ggplotly(SleepvsFormat)
```

## Hypothesis 3: Does the format that I use affect the speed at which I read?

### Hypothesis

$\mu_e =$ Average pages per minute with eink.

$\mu_p =$ Average pages per minute with physical

$H_0: \mu_e = \mu_p$ 

$H_a: \mu_e \neq \mu_p$

```{r}
t.test(PagesPerMinute ~ Format, data = SleepandReadData, alternative = "two.sided")

```

**Conclusion:**

p-value = .562 > $\alpha = .05$, therefore with a $5\%$ significance level, we fail to reject $H_0$. There is not sufficient evidence to support the claim that the format used greatly affects my reading speed.

```{r}

PagesBoxPlot <- ggplot(SleepandReadData, aes(x = Format, y = PagesPerMinute, fill = Format)) +
  geom_boxplot(width = 0.5, 
               outlier.shape = 21, 
               outlier.fill = "white", 
               outlier.color = "black", 
               outlier.size = 2) +
  geom_jitter(width = 0.15, 
              size = 2.5, 
              color = "black", 
              alpha = 0.7) +  
  scale_fill_manual(values = pastel_colors) +
  labs(
    title = "Pages Per Minute by Format",
    y = "Pages Per Minute",
    x = "Format"
  )+
  ylim(0,5.5)


ggplotly(PagesBoxPlot)

```


## Hypothesis 4: Does the format that I use affect how quickly I fall asleep?

### Hypothesis

$\mu_e =$ Average minutes to fall asleep after eink

$\mu_p =$ Average minutes to fall asleep after physical

$H_0: \mu_e = \mu_p$ 

$H_a: \mu_e \neq \mu_p$

```{r}
t.test(MinutesToSleep ~ Format, data = SleepandReadData, alternative = "two.sided")
```

**Conclusion:**

p-value = .3887 > $\alpha = .05$, therefore with a $5\%$ significance level, we fail to reject $H_0$. There is not sufficient evidence to support the claim that the format used greatly affects how long it takes to fall asleep.

```{r}
TimetoSleepPlot <- ggplot(SleepandReadData, aes(x = Format, y = MinutesToSleep, fill = Format)) +
  geom_jitter(width = 0.2, 
              size = 3.5, 
              shape = 21, 
              color = "black", 
              stroke = 1.2) +
  stat_summary(fun = mean, 
               geom = "point", 
               shape = 23, 
               size = 5, 
               fill = "white", 
               color = "black", 
               stroke = 1.2) +
  stat_summary(fun.data = mean_se, 
               geom = "errorbar", 
               width = 0.2, 
               linewidth = 1) +
  scale_fill_manual(values = pastel_colors) +
  labs(
    title = "Minutes to Fall Asleep by Format\n(Note: One outlier above 250 minutes not fully shown)",
    y = "Minutes to Fall Asleep",
    x = "Format"
  ) +
  coord_cartesian(ylim = c(0, 100))  

ggplotly(TimetoSleepPlot)
```

## Additional Graphs

### Stacked/clustered bar chart of average sleep stage minutes by format

```{r}
SleepStageAvg <- SleepandReadData %>%
  group_by(Format) %>%
  summarise(
    REM = mean(REM),
    Light = mean(Light),
    Deep = mean(Deep)
  ) %>%
  pivot_longer(cols = c(REM, Light, Deep), 
               names_to = "Stage", 
               values_to = "AvgMinutes")

StackedBars <- ggplot(SleepStageAvg, aes(x = Format, y = AvgMinutes, fill = Stage)) +
  geom_bar(stat = "identity") +
  scale_fill_manual(values = c("REM" = "#CBBBDD", "Light" = "#BFD8B8", "Deep" = "#FFD8BE")) +
  labs(title = "Average Minutes in Sleep Stages by Format",
       x = "Reading Format",
       y = "Average Minutes",
       fill = "Sleep Stage")+
  theme(legend.position="right")

ggplotly(StackedBars)
```


### Line chart showing reading time by date

```{r}
Lineplot <- ggplot(SleepandReadData, aes(x = Date, y = MinutesReading)) +
  geom_line(linewidth = 1.2, 
            color = "#c3e2c2") +
  geom_point(size = 3, 
             color = "#c3e2c2") +
  labs(title = "Reading Minutes Over Time",
       x = "Date",
       y = "Minutes Reading")

ggplotly(Lineplot)
```

### Bar chart showing average pages read by day of week

```{r}
DayOfWeekSummary <- SleepandReadData %>%
  mutate(DayOfWeek = factor(DayofWeek, 
                            levels = c("Monday", "Tuesday", "Wednesday", "Thursday", "Friday", "Saturday", "Sunday"))) %>%
  group_by(DayOfWeek) %>%
  summarise(AvgPages = mean(Pages, na.rm = TRUE))

DayofWeekPlot <- ggplot(DayOfWeekSummary, aes(x = DayOfWeek, y = AvgPages, fill = DayOfWeek)) +
  geom_col() +
  scale_fill_manual(values = c("#C3E2C2", "#D8C7FF", "#FFD8BE", "#BFD8B8", "#FBE4C9", "#CBBBDD", "#A3C9A8")) +
  labs(title = "Average Pages Read by Day of Week",
       y = "Average Pages",
       x = "Day of Week")

ggplotly(DayofWeekPlot)

```


